---
title: "NeuralNetworks"
author: "Michael Rawat"
date: "04/19/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The codes in this notebook closely follows the original notebook prepared by previous instructors of BIA 6303-Predictive Models: Myles Gartland, Harlan Hays and Xuan Pham. 

## Universal Bank  

You probably have seen the Universal Bank dataset before in BIA 6301: Applied Data Mining. The target variable is PersonalLoan, or whether a customer accepted a loan offer from Universal Bank.

We are going to use the neuralnet package in R to build, train, and validate our neural network models.

```{r}
UB <- read.csv("https://raw.githubusercontent.com/bia6301/BIA6303PredictiveModels/master/Module5/data/UniversalBank.csv")
```

## Data Preprocessing

```{r}
str(UB)
```

```{r}
UB <- UB[,c(-1,-5)]
str(UB)
```


The neuralnet package requires all data to be in numeric format. We cannot have factors. First, we will create dummy variables for the factors.
```{r}
library(dummies)

factors <- UB[,c(6,8:12)]

dummies <- dummy.data.frame(factors)
colnames(dummies)

dummies.keep <- dummies[,c(-2,-4,-6,-8,-10,-12)]
```

Now we need to add these dummy variables together with the numeric variables. 
```{r}
numeric.cols<-UB[,c(1:5,7)] #pull out vars we want to normalize
```

```{r}
df <- cbind(dummies.keep,numeric.cols)
df <- df[,c(3,1,2,4:13)]#moving target variable to the first column
```

Now we can do our train/test split.  

```{r}
library(caret)
trainIndex <- createDataPartition(df$PersonalLoanYes, p = .6, list = FALSE, times = 1)

train <- df[trainIndex,]
test  <- df[-trainIndex,]
```
Neural networks train and converge faster when all variables are normalized. Popular options include min-max normalization (range: 0 to 1) and z-score normalization (mean=0; standard deviation=1).

We have to normalize the numeric columns in train and test separately.


```{r}
normalize<- function(x) #An R-function do carry out min-max normalization
            {
              return(
             (x-min(x))/(max(x)-min(x)))
            }
```

First, we will do min-max normalization on the training set.
```{r}
train.numeric.norm <- normalize(train[,8:13])

train.norm <- cbind(train[,1:7], train.numeric.norm)
```

Second, we repeat the same task for the test set.
```{r}
test.numeric.norm <- normalize(test[,8:13])

test.norm <- cbind(test[,1:7], test.numeric.norm)
```

```{r}
library(neuralnet)
set.seed(123)
clf_nn <- neuralnet(PersonalLoanYes~., data=train.norm, lifesign='full')#be sure to look at the default setting

plot(clf_nn)
```

```{r}
set.seed(123)
clf_nn <- neuralnet(PersonalLoanYes~., data=train.norm, hidden=2, threshold = 0.03, lifesign='full')#be sure to look at the default setting
plot(clf_nn)
```

Now we validate our trained neural network classifier.

```{r}
options(scipen=999)
target.predicted <- compute(clf_nn, test.norm[,-1])

target.predicted.class <- factor(ifelse(target.predicted$net.result>=0.5,"yes","no"))

target.test.class <- factor(ifelse(test$PersonalLoanYes==1,"yes","no"))
```


```{r}
library(caret)
confusionMatrix(target.predicted.class, target.test.class, positive="yes")
```
## More Examples

Check out these blog posts for more examples (including a numeric prediction one): https://www.r-bloggers.com/neuralnet-train-and-test-neural-networks-using-r/ 
https://www.analyticsvidhya.com/blog/2017/09/creating-visualizing-neural-network-in-r/ 